---
title: "Untitled"
author: "Jonathan Bourne"
date: "7 April 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---


Code to provide a quick analysis of the gender paygap data


Where is the glass ceiling thinnest?

Which industries have the most equal balance between men and women at all levels

Which industries are most/least gendered what does this mean for the pay difference?

https://www.ons.gov.uk/employmentandlabourmarket/peopleinwork/employmentandemployeetypes/bulletins/uklabourmarket/march2018
 32.25 milllion people working in the UK

```{r}
basewd <- "/home/jonno/GenderPaygap2"

packages <- c("tidyverse", "readr", "readxl","zoo", "Rtsne", "caret", "ranger" )

new.packages <- packages[!(packages %in% installed.packages()[,"Package"])]
#If packages not installed on computer install them
if(length(new.packages)) install.packages(new.packages)


#Load all packages ready for use
lapply(packages, library, character.only = TRUE)



```


#Load data

```{r}
setwd(basewd)
PayData <- read_csv("UK Gender Pay Gap Data - 2017 to 2018.csv") %>%
  mutate(SicCodes =  gsub("\r|\n", "",SicCodes))

CompSize <- PayData %>%
  group_by(EmployerSize) %>%
  summarise(SizeCounts = n()) %>%
  mutate(AvSize = c(2999.5, 45000, 374.5, 749.5, 12499.5, 225, 225))

PayData <- left_join(PayData, select(CompSize, EmployerSize, AvSize))


SICcodes <- read_csv("SIC07_CH_condensed_list_en.csv")

SIC2 <-read_excel("sic2007summaryofstructurtcm6.xls") %>%
  bind_rows(.,
data_frame(SECTION = "P", X__1 = "EDUCATION",  Division = "1", X__2 = "Education", Group = "1", X__3 = "Trust")
)

SIC2[1,] <- "zzz"

SIC3 <- SIC2

SIC2 <- SIC2%>%
  mutate_all(na.locf)

```


```{r}


PayData %>%
  ggplot(aes(x= DiffMeanHourlyPercent)) +geom_density() +
  coord_cartesian(x= c(-50,100)) + 
  ggtitle("Density plot of mean hourly percent difference") 


PayData %>%
  ggplot(aes(x= EmployerSize, y = DiffMeanHourlyPercent))+ geom_boxplot()

#No real difference by size
PayData %>%
  group_by(EmployerSize) %>%
  summarise(mean = mean(DiffMeanHourlyPercent))

```


#Split by SIC code


```{r}

PayData$SicCodes

#Create 1 hot vector encoding for expertise
Number_of_codes<- PayData$SicCodes %>%
  str_split(., ",") %>% 
  unlist %>% 
  trimws %>%
  as.integer %>%
  as_tibble %>%
  group_by(value) %>%
  summarise(counts = n())

Area <- Number_of_codes$value %>% unique()
 
Area <- Area %>% map(~{
  data_frame(test = grepl(paste0("(^|,)",.x,"($|,)"), PayData$SicCodes))}) %>% #paste0("(^|,)",.x,"($|,)")
  bind_cols() %>%
  setNames(make.names(Area)) 

BroadCats <- SIC3 %>% 
  mutate(X__1 = na.locf(X__1)) %>%
  filter(!is.na(X__2))


SICPay<- Area %>%
  bind_cols(select(PayData, CompanyNumber),.) %>%
  gather(., key = SicCodesTidy, value = value, -CompanyNumber) %>%
  filter(value==1) %>%
  select(-value) %>%
  filter(!is.na(CompanyNumber)) %>%
  left_join(., PayData, "CompanyNumber") %>%
  group_by(CompanyNumber) %>% #just make sure everyone appears the correct amount of times
  mutate(SicCodesTidy = gsub("X", "", SicCodesTidy) %>% as.integer,
         TotCodes = str_count(SicCodes, ",")+1,
         TotOccs = n()) %>%
  ungroup %>%
  #add in sector names
  mutate(chars = substr(SicCodesTidy, start = 1, stop = 2)) %>%
  left_join(select(BroadCats, SECTION:X__2), by=c("chars"="Division")) %>%
  group_by(X__2) %>%
  mutate(MeanMeanSECT = mean(DiffMeanHourlyPercent),
         MeanMedianSECT = median(DiffMedianHourlyPercent)) %>%
  ungroup %>%
  mutate(GenderRatio = (FemaleLowerQuartile+  FemaleLowerMiddleQuartile+ FemaleUpperMiddleQuartile+ FemaleTopQuartile)/4)
  


test <- SICPay %>% 
  group_by(X__2) %>%
  summarise(MeanMeanSECT = first(MeanMeanSECT),
            MeanMedianSECT = first(MeanMedianSECT),
            counts = n(),
            chars = first(chars))

test2 <- SICPay %>% 
  group_by(X__1) %>%
  summarise(MeanMeanSECT = first(MeanMeanSECT),
            MeanMedianSECT = first(MeanMedianSECT),
            counts = n(),
            chars = first(chars))

```


```{r}
test2 <- SICPay %>%
  group_by(SicCodesTidy) %>%
  summarise(counts = n(),
            mean = mean(DiffMeanHourlyPercent))

#most skewed towards women
SICcodes%>% filter(`SIC Code` == 32110)

#second most skewed towards men
SICcodes%>% filter(`SIC Code` == 25720 )


wtf <- SICPay %>%
  filter(SicCodesTidy==47730)
SICcodes%>% filter(`SIC Code` == 47730)

BroadCats <- SIC3 %>% 
  mutate(X__1 = na.locf(X__1)) %>%
  filter(!is.na(X__2))


test4 <- SICPay %>%
  mutate(chars = substr(SicCodesTidy, start = 1, stop = 2)) %>%
  left_join(BroadCats, by=c("chars"="Division")) 



test4 %>%
  ggplot(aes(x = fct_reorder(chars, DiffMeanHourlyPercent, fun = median), y = DiffMeanHourlyPercent )) +
  geom_boxplot() +
  coord_cartesian(y = c(-25,50)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

test4 %>%
  ggplot(aes(x = fct_reorder(X__1, DiffMeanHourlyPercent, fun = median), y = DiffMeanHourlyPercent )) +
  geom_boxplot() +
  coord_cartesian(y = c(-25,50)) 

+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))




test5 <- test4 %>%
  group_by(chars) %>%
  summarise(DiffMeanHourlyPercent = mean(DiffMeanHourlyPercent),
            counts = n(),
            )

```



```{r}
#Convert to matrix for TSNE calulation
TSNEmat <-SICPay %>%
  select(MaleLowerQuartile:FemaleTopQuartile) %>%   
  #select(DiffMeanHourlyPercent:FemaleTopQuartile) %>% 
  as.matrix()

#Some elements are duplicated these need to be removed
TSNEdups <- duplicated(TSNEmat)
sum(TSNEdups)

set.seed(1233)
#Create TSNE embeddings
TsneOut <-TSNEmat[!TSNEdups,] %>%  
  Rtsne(verbose = TRUE)

tsne_plot <- data.frame(x = TsneOut$Y[,1], 
                        y = TsneOut$Y[,2], 
                        Class =SICPay$X__1[!TSNEdups],
                        rank = percent_rank(SICPay$DiffMedianHourlyPercent)[!TSNEdups]) %>%
  mutate(ClassNum = as.factor(Class) %>% as.integer %>% as.character)

ggplot(tsne_plot) + 
  geom_point(aes(x=x, y=y, color= rank>0.5))

tsne_plot %>%
  group_by(Class) %>%
  summarise(ClassNum = first(ClassNum))

```

```{r}
#Convert to matrix for TSNE calulation
TSNEmat2 <-SICPay %>%
  #select(MaleLowerQuartile:FemaleTopQuartile) %>%   
  select(DiffMeanHourlyPercent:FemaleTopQuartile) %>% 
  as.matrix()

#Some elements are duplicated these need to be removed
TSNEdups2 <- duplicated(TSNEmat2)
sum(TSNEdups2)

set.seed(1233)
#Create TSNE embeddings
TsneOut2 <-TSNEmat2[!TSNEdups2,] %>%  
  Rtsne(verbose = TRUE)

tsne_plot2 <- data.frame(x = TsneOut2$Y[,1], 
                        y = TsneOut2$Y[,2], 
                        Class =SICPay$X__1[!TSNEdups2],
                        rank = percent_rank(SICPay$DiffMedianHourlyPercent)[!TSNEdups2]) %>%
  mutate(ClassNum = as.factor(Class) %>% as.integer %>% as.character)

ggplot(tsne_plot2) + 
  geom_point(aes(x=x, y=y, color=ClassNum))

tsne_plot %>%
  group_by(Class) %>%
  summarise(ClassNum = first(ClassNum))
```



```{r}

dataset <- SICPay %>%
  mutate(target = percent_rank(DiffMedianHourlyPercent)) %>%
  select(target, MaleLowerQuartile:FemaleTopQuartile)

dataset <- dataset[!grepl("Female", names(dataset))]

set.seed(192)
trainset <- createDataPartition(dataset$target, p = .75, list = FALSE)


rf_default <- train(y = as.factor(dataset$target[trainset]>0.5), x= dataset[trainset,-1], 
                  method="glm", 
                  family="binomial")


lm_default <- train(y = dataset$target[trainset], x= dataset[trainset,-1], 
                  method="lm")


dataset2 <- dataset %>%
  mutate(target = (target>0.5)*1) %>%
  as.matrix 

rf_ranger_default <- dataset2 %>%
  ranger(data = ., dependent.variable.name = "target", 
                              classification = TRUE)


Outcomes <- predict(rf_ranger_default, data = dataset2[-trainset,], type = "response") 

test <- Outcomes$predictions
confusionMatrix(test, dataset2[-trainset, 1])



Outcomes2 <- predict(rf_default, newdata = dataset[-trainset,])
confusionMatrix(Outcomes2, dataset$target[-trainset]>0.5)

summary(rf_default)

Outcomes2 <- predict(lm_default, newdata = dataset[-trainset,])
postResample(Outcomes2, dataset$target[-trainset])
summary(lm_default)

```

#Promotion Bias

This looks at how the gender balance changes between the quartiles

```{r}
PBias <- PayData %>%
  mutate(ML1 = (MaleLowerMiddleQuartile-MaleLowerQuartile)/(MaleLowerQuartile+1),
         ML2 = (MaleUpperMiddleQuartile-MaleLowerMiddleQuartile)/(MaleLowerMiddleQuartile+1),
         ML3 = (MaleTopQuartile-MaleUpperMiddleQuartile)/(MaleUpperMiddleQuartile+1),
         FL1 = (FemaleLowerMiddleQuartile-FemaleLowerQuartile)/(FemaleLowerQuartile+1),
         FL2 = (FemaleUpperMiddleQuartile-FemaleLowerMiddleQuartile)/(FemaleLowerMiddleQuartile+1),
         FL3 = (FemaleTopQuartile-FemaleUpperMiddleQuartile)/(FemaleUpperMiddleQuartile+1),
         MBias = (ML1+ML2+ML3)/3,
         FBias = (FL1+FL2+FL3)/3,
         # RatQ2 = (MaleLowerMiddleQuartile*(1-MaleLowerQuartile)+1)/(MaleLowerQuartile*(1-MaleLowerMiddleQuartile)+1),
         # RatQ3 = (MaleUpperMiddleQuartile*(1-MaleLowerMiddleQuartile)+1)/(MaleLowerMiddleQuartile*(1-MaleUpperMiddleQuartile)+1),
         # RatQ4 = (MaleTopQuartile*(1-MaleUpperMiddleQuartile)+1)/(MaleUpperMiddleQuartile*(1-MaleTopQuartile)+1),
         RatQ2 = (MaleLowerMiddleQuartile*(1-MaleLowerQuartile))/(MaleLowerQuartile*(1-MaleLowerMiddleQuartile)),
         RatQ3 = (MaleUpperMiddleQuartile*(1-MaleLowerMiddleQuartile))/(MaleLowerMiddleQuartile*(1-MaleUpperMiddleQuartile)),
         RatQ4 = (MaleTopQuartile*(1-MaleUpperMiddleQuartile))/(MaleUpperMiddleQuartile*(1-MaleTopQuartile)),
         MeanRat = (RatQ2+RatQ3+RatQ4)/3,
         Percmale = (MaleLowerQuartile + MaleLowerMiddleQuartile + MaleUpperMiddleQuartile + MaleTopQuartile)/4)



MaleCenter <- 1:nrow(PBias) %>%
  map_dbl(~{
    PBias[.x,] %>% select(MaleLowerQuartile, MaleLowerMiddleQuartile, MaleUpperMiddleQuartile, MaleTopQuartile) %>%t %>%
      weighted.mean(1:4, .)
  })

FemaleCenter <- 1:nrow(PBias) %>%
  map_dbl(~{
    PBias[.x,] %>% select(FemaleLowerQuartile, FemaleLowerMiddleQuartile, FemaleUpperMiddleQuartile, FemaleTopQuartile) %>%t %>%
      weighted.mean(1:4, .)
  })


PBias <- PBias %>%
  mutate(MaleCenter = MaleCenter,
         FemaleCenter = FemaleCenter)

mean(PBias$MBias, na.rm = T)
weighted.mean(PBias$MBias, PBias$AvSize, na.rm = T)

mean(PBias$FBias, na.rm = T)
weighted.mean(PBias$FBias, PBias$AvSize, na.rm = T)


t.test(PBias$MBias, PBias$FBias, paired = T)

PBias %>%
  select(MBias, FBias, AvSize) %>%
  gather(., key = key, value = value, -AvSize) %>%
  ggplot(aes(x= value, colour = key)) + geom_density()+
  coord_cartesian(x=c(-0.7,0.8)) +
  labs(x = "Mean promotion bias", title = "Internal Promotion bias")

PBias %>%
  select(MBias, FBias) %>%
  ggplot(aes(x= MBias, y = FBias)) + geom_point()+
  coord_cartesian(x=c(-0.5,1.5), y=c(-0.7,1.5))

PBias %>%  coord_cartesian(x=c(0.9,1.1))
  select(ML1:FL3) %>%
  gather %>%
  ggplot(aes(colour= key, x = value)) + geom_density() +
   coord_cartesian(x=c(-1.5,0.5))


#The bias for going from Q4 to Q3 shows that

PBias %>%
  select(ML1:FL3) %>%
  gather %>%
  group_by(key) %>%
  summarise(mean = mean(value, na.rm = T),
            median = median(value, na.rm = T))

#percent of workforce covered
sum(PBias$AvSize)/32250000

TotalLength <- 200000

#creates a vector that replicates a companies bias a number of time proportional to its weight in society

setwd(basewd)
if(!file.exists("WeightedData.rmd")){
WeightedOut <- c("CompanyNumber","ML1", "ML2","ML3","FL1","FL2","FL3","MBias","FBias", "MeanRat", "MaleCenter", "FemaleCenter" )  %>% 
 map(~{
   print(.x)
   target <- select(PBias, .x)
test <-  map2_df(.x = 1:nrow(PBias), .y =PBias$AvSize/sum(PBias$AvSize), ~{

  z<- data.frame(value = rep(target[.x,1], times = floor(.y*TotalLength)
                             ) %>% unlist
                 ) %>%
      set_names(names(target))
    
    return(z)
  }) 
print(class(test))
return(test)

})  %>%
 bind_cols()

saveRDS(WeightedOut , "WeightedData.rmd")
} else{
  
  readRDS("WeightedData.rmd")
}

 WeightedOut %>%
  select(MBias, FBias) %>%
  gather %>% 
    ggplot(aes(colour= key, x = value)) + geom_density() +
   coord_cartesian(x=c(-0.6,0.8))

WeightedOut  %>%
  gather %>%
  group_by(key) %>%
  summarise(mean = mean(value, na.rm = T),
            median = median(value, na.rm = T))

 WeightedOut %>%
  select(ML1:FL3) %>%
  gather %>% 
    ggplot(aes(colour= key, x = value)) + geom_density() +
   coord_cartesian(x=c(-1.5,0.8))
 
```


#test bias
```{r}
x1<- seq(1,100,1)
x2<- seq(100,0.1,-1)

x1<- seq(.1,1,.01)
x2<- seq(1,0.1,-.01)

x1<- c(0,0,0)
x2<- c()

Out <-(x2*(1-x1))/(x1*(1-x2))
plot(Out[80:100])

PBias2 <- PayData %>%
  mutate(Q1 =MaleLowerQuartile/100,
         Q2 = MaleLowerMiddleQuartile/100,
         Q3 = MaleUpperMiddleQuartile/100,
         Q4 = MaleTopQuartile,
         # RatQ2 = (MaleLowerMiddleQuartile*(1-MaleLowerQuartile)+1)/(MaleLowerQuartile*(1-MaleLowerMiddleQuartile)+1),
         # RatQ3 = (MaleUpperMiddleQuartile*(1-MaleLowerMiddleQuartile)+1)/(MaleLowerMiddleQuartile*(1-MaleUpperMiddleQuartile)+1),
         # RatQ4 = (MaleTopQuartile*(1-MaleUpperMiddleQuartile)+1)/(MaleUpperMiddleQuartile*(1-MaleTopQuartile)+1),
         RatQ2 = (Q2*(1-Q1)+1)/(Q1*(1-Q2)+1),
         RatQ3 = (Q3*(1-Q2)+1)/(Q2*(1-Q3)+1),
         RatQ4 = (Q4*(1-Q3)+1)/(Q3*(1-Q4)+1),
         RatioQ2 = (Q2*(1-Q1))/(Q1*(1-Q2)),
         RatioQ3 = (Q3*(1-Q2))/(Q2*(1-Q3)),
         RatioQ4 = (Q4*(1-Q3))/(Q3*(1-Q4)),
         MeanRat = (RatQ2+RatQ3+RatQ4)/3,
         MeanRat2 = (RatioQ2+RatioQ3+RatioQ4),
         Percmale = (MaleLowerQuartile + MaleLowerMiddleQuartile + MaleUpperMiddleQuartile + MaleTopQuartile)/4) %>%
  select(RatQ2:MeanRat2) %>%
    filter(MeanRat2<25, MeanRat2>-25) 


PBias2  %>%
    filter(MeanRat2<25, MeanRat2>-25) %>% 
  ggplot(aes(x= MeanRat2)) +geom_density()



median(PBias2$MeanRat2, na.rm = T)

```



```{r}

PBias %>%
  ggplot(aes(x= MeanRat)) + geom_density()+
  coord_cartesian(x=c(0.9,1.05))


PBias %>%
  ggplot(aes(x= RatQ2)) + geom_density()+
  coord_cartesian(x=c(0.94,1.1))

PBias %>%
  ggplot(aes(x= Maleweight, y = Femaleweight)) +
  geom_point()

cor(PBias$DiffMedianHourlyPercent, PBias$MaleCenter)

 WeightedOut%>%
  select(FemaleCenter, MaleCenter) %>%
  gather %>% 
    ggplot(aes(colour= key, x = value)) + geom_density()

 mean(PBias$MaleCenter, na.rm = T)
weighted.mean(PBias$MaleCenter, PBias$AvSize, na.rm = T)

mean(PBias$FemaleCenter, na.rm = T)
weighted.mean(PBias$FemaleCenter, PBias$AvSize, na.rm = T)


t.test(PBias$MaleCenter, PBias$FemaleCenter)
 
```



```{r}

dataset <- PBias %>%
  mutate(target = percent_rank(DiffMedianHourlyPercent)) %>%
  select(target, MeanRat, MaleCenter, FemaleCenter) %>%
  filter(is.finite(MeanRat), is.finite(FemaleCenter))

set.seed(192)
trainset <- createDataPartition(dataset$target, p = .75, list = FALSE)



lm_default <- train(y = dataset$target[trainset], x= dataset[trainset,-1], 
                  method="lm")

summary(lm_default)

Outcomes2 <- predict(lm_default, newdata = dataset[-trainset,])
postResample(Outcomes2, dataset$target[-trainset])

```



```{r}

test <- WeightedOut %>%
  left_join(select(SICPay, CompanyNumber, SicCodesTidy, X__1, X__2, DiffMedianHourlyPercent, GenderRatio)) %>%
  as_tibble()%>%
  filter(is.finite(MeanRat)) %>%
  filter(!is.na(X__1))


test %>%
  ggplot(aes(x =fct_reorder( X__1, MeanRat), y = MeanRat)) + geom_boxplot() +
  coord_cartesian(y = c(0.8,1.3))

test2 <- test %>%
  filter(is.finite(MeanRat)) %>%
  group_by(X__1) %>%
    summarise(Rat = median(MeanRat),
              MaleCenter = median(MaleCenter),
              FemaleCenter = median(FemaleCenter),
              DiffMedianHourlyPercent = median(DiffMedianHourlyPercent),
              counts = n(),
              GenderRatio = median(GenderRatio)) %>%
  arrange(Rat)

test2

cor(test2$Rat, test2$MaleCenter)
cor(test$MeanRat, test$MaleCenter)

cor(test2$Rat, test2$FemaleCenter)
cor(test$MeanRat, test$FemaleCenter)

cor(test$FemaleCenter, test$MaleCenter)

cor(test2$GenderRatio, test2$Rat)
cor(test$GenderRatio, test$MeanRat)

cor(test2$GenderRatio, test2$FemaleCenter)
cor(test$GenderRatio, test$FemaleCenter)



```



#Sector Sigmoid fit 

This chunk shows how the relationship between the promotions ratio and the pay gap follows a sigmoid shape.

It shows that if you are in an industry with better promotion prospects you are also in an industry with a smaller pay gap

```{r}
fit <- nls(y ~ SSlogis(x, Asym, xmid, scal), data = data.frame( x= test2$Rat, y = test2$DiffMedianHourlyPercent) %>% filter(y !=0))

summary(fit)

lines(seq(0.5, 4, length.out = 100), 
      predict(fit, newdata = data.frame(x = seq(0.5, 4, length.out = 100))))

SigmoidCurve <- data_frame( y = predict(fit, newdata = data.frame(x = seq(1, max(test2$Rat), length.out = 100)))[1:100],
                   x = seq(1, max(test2$Rat), length.out = 100))

test2 %>%
  ggplot() +geom_point(aes(x= Rat, y = DiffMedianHourlyPercent)) +
  geom_line(data = SigmoidCurve, aes(x= x, y = y), colour = "red") +
  labs(x= "Male Promotion Liklihood", 
       y ="Percent Difference in Median Hourly Wage", 
       title = "Sector Relationship Between promotion likelyhood and Pay gap")

```

